Data Pre-Processing

### 2021-05-14 AST-Framing-Ontario/Data Pre-Processesing.rmd
Changes From Past Week
- Parameters at the top
- Function for ligature replacement
- Double column function
- Other code clean up and bug fixes

---
title: "Data Pre-Processing"
author: "Jason Lam"
date: "11/03/2021"
output:
  pdf_document: default
  word_document: default
---

@@

Load the following packages:
```{r}
library(tm)
library(readr)
library(tools)
library(dplyr)
library(tidytext)
library(topicmodels)
library(ggplot2)
library(wordcloud)
library(word2vec)
library(pdftools)
library(reshape2)
library(text2vec)
```

@@

@@

```{r}
pdfLocation <- "Texts/Academic Papers NoRef"

@@

# use of `sep = ""` argument is important to prevent the addition of spaces, which is the default for the paste function 
newFile <- paste(txtLocation, "/", file_path_sans_ext(basename(address)), ".txt", sep = "")
write(test, file = newFile)
print( paste(file_path_sans_ext(basename(address)), ".txt Has been created", sep = ""))

# Creates uniquely id text store in 'Environment'
assign(paste("text", n, sep = ""), readChar(newFile, file.info(newFile)$size))
n <- n+1
}
```

@@

for (i in seq(1, n-1, 1))
{
text[i] <- eval(parse(text = paste("text", i, sep = "")))
}

# Turns vector into data frame 
textDF <- data.frame(doc_id, text)
```

@@

Next, convert the data frame to a corpus:
```{r}
df_source <- DataframeSource(textDF)
corpus <- VCorpus(df_source)
# Example 
content(corpus[[1]])
```

@@

NAwords <- c("school", "transportation", "can", "use", "active", "et", "al", "travel", "study", "elsevier", "na", "sciencedirect", "journal", "transport", "mode", "modes", "children", "the", "use", "n", "s", "schools", "also", "p", "b", "used", "number")
corpus <- tm_map(corpus, removeWords, NAwords)
corpus <- tm_map(corpus, removeNumbers)

@@

# Example
content(corpus[[1]])
```

@@

@@

safe = word_vectors["wellbeing", , drop = FALSE] - word_vectors["physical", , drop = FALSE] + word_vectors["mental", , drop = FALSE]

cos_sim = sim2(x = word_vectors, y = safe, method = "cosine")

head(sort(cos_sim[,1], decreasing = TRUE), 10)
```

@@

'n' represents the number of subfolders and subsequently corpora that you intent the produce. 'corpusKey' will indicate the numbering convention of the corpora and analysis in subsequent code. 
```{r}
# n - number of corpora
n <- 4
corpusNum <- seq(1, n, 1)
corpusKey <- data.frame(corpusNum, subFolders)
```

@@

corpusNum <- c()

@@

for (i in seq(1, n-1, 1))
{
text[i] <- eval(parse(text = paste("text", i, sep = "")))
}

@@

corpusKey <- data.frame(corpusNum, subFolders)

rm(temp_df_source, tempCorpus, address, doc_id, fileBaseNames, filePath, i, m, n, newFile, pdfLocation, subFolder, test, txtLocation)
```

@@

```{r}
for (i in seq(1, n, 1))
{
  tempDF <- eval(parse(text = paste("cleanTextDF", i, sep = "")))
  it_train = itoken(tempDF$text, progressbar = FALSE)
  assign(paste("vocab", i, sep= ""), create_vocabulary(it_train))
  tempVocab <- create_vocabulary(it_train)
  assign(paste("vocab", i, sep= ""), prune_vocabulary(tempVocab, term_count_min = 5))
}

@@

paste(sort(NAwords4), collapse = " ")

@@

>>>>

---
title: "Data Pre-Processing"
author: "Jason Lam"
date: "11/03/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
params:
  Academic_Papers: "Texts/Academic Papers NoRef"
  Academic_Papers_DoubCol: "Texts/Academic Papers NoRef DoubleCol"
  Consortia: "Texts/Consortia"
  Municipalities_Regions: "Texts/Municipalities or Regions"
  School Boards: "Texts/School Boards"
  AP_stopW: c("school", "transportation", "can", "use", "active", "et", "al", "travel", "study", "elsevier", "na", "sciencedirect", "journal", "transport", "mode", "modes", "children", "the", "n", "s", "schools", "also", "p", "b", "used", "number", "data", "research", "variable", "variables", "author", "authors", "researcher", "researchers", "project", "instrument", "may", "model", "studies", "one", "table", "results", "found", "likely", "using", "less", "high", "within", "associated", "however", "among", "two", "km", "eg", "m", "j", "c", "sample", "per", "three", "r", "ie", "across", "around", "well", "a", "example", "although", "t", "fig", "signiufbcant", "1/4", "ats", "d", "literature", "overall", "'", "f", "paper", "e", "therefore", "thus", "methods", "deparmant", "result", "collected", "-")
  AP_DoubCol_stopW: c("school", "transportation", "can", "use", "active", "et", "al", "travel", "study", "elsevier", "na", "sciencedirect", "journal", "transport", "mode", "modes", "children", "the", "n", "s", "schools", "also", "p", "b", "used", "number", "data", "research", "variable", "variables", "author", "authors", "researcher", "researchers", "project", "instrument", "may", "model", "studies", "one", "table", "results", "found", "likely", "using", "less", "high", "within", "associated", "however", "among", "two", "km", "eg", "m", "j", "c", "sample", "per", "three", "r", "ie", "across", "around", "well", "a", "example", "although", "t", "fig", "signiufbcant", "1/4", "ats", "d", "literature", "overall", "'", "f", "paper", "e", "therefore", "thus", "methods", "deparmant", "result", "collected", "-")
  Con_stopW: c("login", "css", "sir", "waterloo", "niagara", "john", "school", "transportation", "york", "ottawa", "can", "london", "provides", "elgin", "find", "make", "lord", "holy", "toronto", "get", "please", "will", "also", "copyright", "canada", "ontario", "translate", "download", "email", "one", "arthur", "de", "info", "website", "x")
  Mun_Reg_stopW: c("york", "lane", "school", "transportation", "region", "halton", "can", "use", "guelph", "kingston", "durham", "get", "uf", "ontario", "contact", "also", "covid", "c", "e", "one", "eg", "-", "please", "canada", "n", "page", "canadian", "email", "o", "peel", "fort", "frances", "phone", "r", "ottawa", "barrie", "km", "next", "per", "thunder", "simcoe", "website", "wellington", "andor", "catholic", "k", "m", "middlesex", "oxford", "perth", "po", "provincial", "tollfree", "toronto", "ufc")
  SB_stopW: c("and", "asst", "board", "can", "cancellations", "catholic", "contact", "district", "east", "gta", "https", "login", "may", "or", "please", "school", "schools", "stp", "student", "students", "the", "transportation", "tri", "waterloo", "website", "will", "uf", "ontario", "-", "us", "north", "km", "county", "ottawa", "guelph", "ext", "phone", "e", "ps", "'s", "du", "toronto", "email", "o", "ottschoolbus", "simcoe", "canada", "careers", "covid", "facts", "ufd", "york", "pm", "ufa", "ufe", "wrdsb", "andor", "erin", "kg", "transporta", "cambridge", "httpsvimeocom", "muskoka", "stopr", "ufc", "windsor", "a", "durham", "geoquery", "hamiltonwentworth", "httpswwwwrdsbcaplanningwp", "hwdsb", "ic", "jk", "kawartha", "ng", "northeastern", "sudbury", "timiskaming", "wellingtondu")

---

@@

Load the following packages:
```{r}
library(dplyr)
library(ggraph)
library(ggplot2)
library(igraph)
library(pdftools)
library(readr)
library(reshape2)
library(stringr)
library(text2vec)
library(tidyr)
library(tidytext)
library(tm) 
library(tools)
library(topicmodels)
library(word2vec)
library(wordcloud)
```

@@

## Function to clean text of Ligatures

The function below allows you to find and replace certain strings. The main use in this context is to replace encoding errors where specific sequences of character such as "fi" or "fl" are encoded as ligatures in the pdf. During the conversion process those character can be lost and their subsequent words are misintrupted.

https://stackoverflow.com/questions/17517319/r-replacing-foreign-characters-in-a-string 
```{r}
to.plain <- function(s) {

   # 1 character substitutions
   old1 <- "šžþàáâãäåçèéêëìíîïðñòóôõöùúûüý"
   new1 <- "szyaaaaaaceeeeiiiidnooooouuuuy"
   s1 <- chartr(old1, new1, s)

   # 2 character substitutions
   old2 <- c("<U+FB03>", "<U+FB01>", "<U+FB00>", "<U+FB02>")
   new2 <- c("ffi", "fi", "ff", "fl")
   s2 <- s1
   for(i in seq_along(old2)) s2 <- gsub(old2[i], new2[i], s2, fixed = TRUE)

   s2
}
```

@@

```{r}
pdfLocation <- params$Academic_Papers

@@

# use of `sep = ""` argument is important to prevent the addition of spaces, which is the default for the paste function 
newFile <- paste(txtLocation, "/", file_path_sans_ext(basename(address)), ".txt", sep = "")
print( paste(file_path_sans_ext(basename(address)), ".txt Has been created", sep = ""))

write(test, file = newFile)

# Creates uniquely id text store in 'Environment'
assign(paste("text", n, sep = ""), readChar(newFile, file.info(newFile)$size))

n <- n+1
}
```

@@

for (i in seq(1, n-1, 1))
{
  #
text[i] <- to.plain(eval(parse(text = paste("text", i, sep = ""))))
}

# Turns vector into data frame 
textDF <- data.frame(doc_id, text)

@@

Next, convert the data frame to a corpus:
```{r}
df_source <- DataframeSource(textDF)
corpus <- VCorpus(df_source)
# Example 
content(corpus[[2]])
```

@@

NAwords <- c("login", "css", "sir", "waterloo", "niagara", "john", "school", "transportation", "york", "ottawa", "can", "london", "provides", "elgin", "find", "make", "lord", "holy", "toronto", "get", "please", "will", "also", "copyright", "canada", "ontario", "translate", "download", "email", "one", "arthur", "de", "info", "website", "x")
corpus <- tm_map(corpus, removeWords, NAwords)
corpus <- tm_map(corpus, removeNumbers)

@@

# Example
content(corpus[[2]])
```

@@

Bi-gram visualizations 
```{r}
bigrams_separated <- test_bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

bigram_counts <- bigrams_separated %>% 
  count(word1, word2, sort = TRUE)

# Adjust the parameter below to filter out based on number of occurence
# Error will occur if n is outide the range of occurences 
bigram_graph <- bigram_counts %>%
  filter(n > 5) %>%
  graph_from_data_frame()

set.seed(2021)

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1)

@@

#safe = word_vectors["wellbeing", , drop = FALSE] - word_vectors["physical", , drop = FALSE] + word_vectors["mental", , drop = FALSE]

#cos_sim = sim2(x = word_vectors, y = safe, method = "cosine")

#head(sort(cos_sim[,1], decreasing = TRUE), 10)
```

@@

## Function to clean text of Ligatures
```{r}
to.plain <- function(s) {

   # 1 character substitutions
   old1 <- "šžþàáâãäåçèéêëìíîïðñòóôõöùúûüý"
   new1 <- "szyaaaaaaceeeeiiiidnooooouuuuy"
   s1 <- chartr(old1, new1, s)

   # 2 character substitutions
   old2 <- c("<U+FB03>", "<U+FB01>", "<U+FB00>", "<U+FB02>")
   new2 <- c("ffi", "fi", "ff", "fl")
   s2 <- s1
   for(i in seq_along(old2)) s2 <- gsub(old2[i], new2[i], s2, fixed = TRUE)

   s2
}
```

'n' represents the number of subfolders and subsequently corpora that you intent the produce. 'corpusKey' will indicate the numbering convention of the corpora and analysis in subsequent code. 
```{r}
# n - number of corpora
n <- 5
corpusNum <- seq(1, n, 1)
corpusKey <- data.frame(corpusNum, subFolders)
```

@@

NAwords5 <- c()

@@

for (i in seq(1, n-1, 1))
{
text[i] <- to.plain(eval(parse(text = paste("text", i, sep = ""))))
}

@@

rm(temp_df_source, tempCorpus, address, doc_id, fileBaseNames, filePath, i, m, newFile, pdfLocation, subFolder, test, txtLocation)
```

@@

```{r}
n <- 5 
for (i in seq(1, n, 1))
{
  tempDF <- eval(parse(text = paste("cleanTextDF", i, sep = "")))
  it_train = itoken(tempDF$text, progressbar = FALSE)
  assign(paste("vocab", i, sep= ""), create_vocabulary(it_train))
  tempVocab <- create_vocabulary(it_train)
  assign(paste("vocab", i, sep= ""), prune_vocabulary(tempVocab, term_count_min = 5))
}

@@

paste(sort(c("tm", "readr", "tools", "dplyr", "tidytext", "topicmodels", "ggplot2", "wordcloud", "word2vec", "pdftools", "reshape2", "text2vec", "stringr", "tidyr", "igraph")), collapse = " ")

@@

# Coverts double coloumn documents to single column

The first chuck sets up the function used in the second section.

https://stackoverflow.com/questions/42541849/extract-text-from-two-column-pdf-with-r 
```{r}

trim <- function (x) gsub("^\\s+|\\s+$", "", x)

QTD_COLUMNS <- 2
read_text <- function(text) {
  result <- ''
  #Get all index of " " from page.
  lstops <- gregexpr(pattern =" ",text)
  #Puts the index of the most frequents ' ' in a vector.
  stops <- as.integer(names(sort(table(unlist(lstops)),decreasing=TRUE)[1:2]))
  #Slice based in the specified number of colums (this can be improved)
  for(i in seq(1, QTD_COLUMNS, by=1))
  {
    temp_result <- sapply(text, function(x){
      start <- 1
      stop <-stops[i] 
      if(i > 1)            
        start <- stops[i-1] + 1
      if(i == QTD_COLUMNS)#last column, read until end.
        stop <- nchar(x)+1
      substr(x, start=start, stop=stop)
    }, USE.NAMES=FALSE)
    temp_result <- trim(temp_result)
    result <- append(result, temp_result)
  }
  result
}

# Create a loop of the code below to run through list of documents

pdfLocation <- "Texts/Academic Papers NoRef DoubleCol"

 if (file.exists(paste(pdfLocation, "txt"))){
    txtLocation <- paste(pdfLocation, "txt")
  }else{
    dir.create(paste(pdfLocation, "txt"))
    txtLocation <- paste(pdfLocation, "txt")
  }
  fileBaseNames <- list.files(pdfLocation, pattern = "*.pdf", full.names=FALSE)
  filePath <- list.files(pdfLocation, pattern = "*.pdf", full.names=TRUE)

for (pdf in filePath)
  {
txt <- pdf_text(pdf)
result <- ''
for (i in 1:length(txt)) { 
  page <- txt[i]
  t1 <- unlist(strsplit(page, "\n"))      
  maxSize <- max(nchar(t1))
  t1 <- paste0(t1,strrep(" ", maxSize-nchar(t1)))
  result = append(result,read_text(t1))
}
newFile <- paste(txtLocation, "/", file_path_sans_ext(basename(pdf)), ".txt", sep = "")
write(result, newFile)
}

```{r}
rm(list=ls(pattern = "^text"))
```
